Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-09-20 paper/main.tex
replace nelder-mead with l-bfgs-b

We chose to optimize for the damping of agents' movement within the medium, 3 interaction parameters
of the Mie potential $n,m,V_0$ a single radius for all agents and growth rates which were chosen
individually for each cell.
We removed any stochastic motion contributions by setting $D=0$.
This leads to a total of $5+1\times 6=11$ parameters.
To estimate these parameters we used the differential evolution~\cite{Storn1997} global optimization
algorithm which was initially seeded with a sample generated by the latin-hypercube~\cite{McKay1979}
method.
Finally, we performed a local minimization using the approach by Nelder-Mead~\cite{Powell1973}.

@@

Since we have already optimized said system, we can reduce the computational demand by choosing the
parameters which were obtained by our previous attempts and extend them with the parameters that the
daughter cells will carry.
In addition to assigning growth rates individually, but we also need to consider division lengths at
which the bacteria split.
Since our simulation involves the division of 4 cells into 8 new ones, in principle, this would lead
us to a total of $8$ new growth rates and $4$ division lengths.
But to decrease the size of the parameter space, we assume that the growth rates of the daughter
cells are identical between every sibling, leaving us at only $4$ new growth rates and a total of
$8$ parameters.
Future approaches should aim to loosen this assumption, allowing to optimize every new growth rate.
To optimize this system, we again used the same optimization approach by seed with samples generated
by latin-hypercube~\cite{McKay1979}, then applying the differential evolution
algorithm~\cite{Storn1997} and finally a local minimization with Nelder-Mead~\cite{Powell1973}.
We used the value $0.5$ for the parent penalty, which aims to provide a smooth transition between
not accounting for them and assigning full weighting to this difference.
Figure~\ref{fig:celldiffs-progression-with-division} shows the difference between the cell mask as
given by the data and the cell mask predicted by the best fit of our model.

>>>>

We chose to optimize for the damping of agents' movement within the medium, 3 interaction parameters
of the Mie potential $n,m,V_0$ a single radius for all agents and growth rates which were chosen
individually for each cell.
We removed any stochastic motion contributions by setting $D=0$.
This leads to a total of $5+1\times 6=11$ parameters.
To estimate these parameters we used the differential evolution~\cite{Storn1997} global optimization
algorithm which was initially seeded with a sample generated by the latin-hypercube~\cite{McKay1979}
method.
Finally, we performed a local minimization using the L-BFGS-B algorithm~\cite{Liu1989}.

@@

Since we have already optimized said system, we can reduce the computational demand by choosing the
parameters which were obtained by our previous attempts and extend them with the parameters that the
daughter cells will carry.
In addition to assigning growth rates individually, but we also need to consider division lengths at
which the bacteria split.
Since our simulation involves the division of 4 cells into 8 new ones, in principle, this would lead
us to a total of $8$ new growth rates and $4$ division lengths.
But to decrease the size of the parameter space, we assume that the growth rates of the daughter
cells are identical between every sibling, leaving us at only $4$ new growth rates and a total of
$8$ parameters.
Future approaches should aim to loosen this assumption, allowing to optimize every new growth rate.
To optimize this system, we again used the same optimization approach by seed with samples generated
by latin-hypercube~\cite{McKay1979}, then applying the differential evolution
algorithm~\cite{Storn1997} and finally a local minimization via L-BFGS-B~\cite{Liu1989}.
We used the value $0.5$ for the parent penalty, which aims to provide a smooth transition between
not accounting for them and assigning full weighting to this difference.
Figure~\ref{fig:celldiffs-progression-with-division} shows the difference between the cell mask as
given by the data and the cell mask predicted by the best fit of our model.
